Cover letter for the IJRR submission \\ ``An efficient acyclic contact planner for multiped robots''

### 2017-01-20 body/appendix_optim.tex
Review appendix C

\section{Addressing $\mathcal{P}_3$}
\label{app:optim}
The stair climbing and the standing up scenarios were validated with the trajectory optimization scheme provided in \citeauthor{Carpentier2016}. 
The optimization is formulated as an optimal control problem that aims at generating a valid trajectory for the center of mass.
To address the others scenarios, we propose a new implementation, entirely open source (\url{https://github.com/stonneau/python_sandbox}), which can be integrated directly in our motion planner software. This formulation uses the center of mass acceleration and angular momentum as input variables, while previously the cotnact forces were used.
The center of mass trajectory resulting from the optimization is then turned into a collision-free whole-body trajectory.

@@

\begin{align} \label{eq:new_eul_acc}
\mathbf{G} \bm{\beta} = 
\underbrace{\mat{m (\mathbf{\ddot{c}} - \mathbf{g}) \\m (\mathbf{\ddot{c}} - \mathbf{g}) + \dot{L}}}_{\mathbf{w}}
%~ \underbrace{m (\mathbf{\ddot{c}} - \mathbf{g})  \\ m (\mathbf{\ddot{dc}} - \mathbf{g}) + \mathbf{\dot{L}} }_{\mathbf{w}}
\end{align}

where $\dot{L}$ is the angular momentum expressed at the com.
Eq.~\ref{eq:new_eul_acc} defines a 6 dimensional cone $\mathcal{K}$ ~\citep{qiu:dhm:2011,Caron2015}. For a given set of contacts,
this cone determines all the contact forces that generate an admissible wrench $\mathbf{w}$ such that the robot won't fall.
The face form of $\mathcal{K}$ can be computed using the double description method~\citep{Fukuda1996}, resulting in the following linear inequalities:

@@

The objective 
is then to plan a trajectory for the center of mass such that the generated $\mathbf{w}$ always verifies Eq.~\ref{eq:cone_k}. 
We now consider two contact configurations  $\mathbf{q}_0$ and $\mathbf{q}_1$ computed by our planner: in the general case one contact is broken and one created to get from
$\mathbf{q}_0$ to $\mathbf{q}_1$. The motion is thus constrained by three cones $\mathcal{K}_s, s =0 \dots 2$. We manually define the duration of each of the three contact phases.
We call the total duration of the motion $T$, and formulate the following optimization problem:

\begin{equation} \label{eq:lin_prog_rob} \begin{aligned}
%~ \find \quad & \mx{R} \in \Rv{3\times T}, \mxdd{C} \in  \Rv{6\times T} \\
\minimize_{\mathbf{c}(t), \mathbf{w}(t)}  \quad & \sum_{s=0}^{2} \int_{t_s}^{t_s + \Delta t_s}  h(\mathbf{c}, \mathbf{w})  \\
\st \quad & \mathbf{\dot{c}} = f(\mathbf{c}, \mathbf{w}) \\
	\quad & \mathbf{A}_s \mathbf{w}(t)  \leq \mathbf{b}_s , \forall t \in  \left[t_s, t_s + \Delta t_s \right[ \\
	\quad & \mathbf{D}_s \mathbf{c}(t)  \leq \mathbf{d}_s , \forall t \in  \left[t_s, t_s + \Delta t_s \right[ \\
	\quad & \mathbf{c}(0)  = \mathbf{c}_{\mathbf{q}_0} \\
	\quad & \mathbf{c}(T)  = \mathbf{c}_{\mathbf{q}_1} \\
	\quad & \mathbf{c}(0)  = \mathbf{\dot{c}}(0)  = \mathbf{\ddot{c}}(0) = \mathbf{0}\\
	\quad & \mathbf{c}(T)  = \mathbf{\dot{c}}(T)  = \mathbf{\ddot{c}}(T) = \mathbf{0}\\
    %~ \quad & ({\vct{r}}\tra\vct{r}=1,  \forall t  ?) \\
    %~ \quad & \delta t \sum\limits_1^{T-1}{\vc{r}^1}\tra {\vc{r}^1} \ge t_{transit} \\
    %~ \quad & \mx{A}_{\int} \mxdd{C} \leq \vc{a}_{\int} \\
    %~ \quad & \mx{V}_{\int} \mxd{C} \leq \vc{v}_{\int} \\
    %~ \quad & \mx{B}_{\int}^{1} \mx{C} \leq \vc{b}_{\int} \\
    %~ \quad & [\vc{c}_0\tra, \vcd{c}_0\tra]\tra = \vc{x}_0 \\
    %~ \quad & \vc{c}_T = \vc{c}_T^{input} \\
\end{aligned} \end{equation}

As a cost function, $h$ typically minimizes the angular momentum and acceleration variation over the whole trajectory. $f$ is the integration method
that ties together the position, velocity and acceleration of the center of mass. 
$\mathbf{D}_s$ and $\mathbf{d}_s$ denote stacked kinematic constraints on the center of mass position, determined by the active contact locations.
The inequalities for each contact are determined in the same way that the reachable workspace is computed in Appendix~\ref{app:rom}, with
the effector serving as root (a triangulated object is a polytope, and as such can be described as a set of linear inequalities).

@@

As an output, the solution of this optimization problem gives us a admissible center of mass trajectory.
To compute the whole body motion, we use a two step approach.

First, we plan a kinematic motion for the robot, subject to the contact constraints. We also constrain the center of mass to follow
the computed trajectory. This is achieved using a constraint-based RRT planner~\citep{7759083}. As a result we obtain a whole-body motion that is collision free.

@@

This motion is then validated in our dynamic simulator~\citep{DelPrete2015b}, using the trajectory as a reference postural task to an inverse dynamics
solver, that accounts for the torque constraints of the robot. This final result is shown in the videos.

>>>>

% !TEX root =  ../main.tex
\section{Addressing $\mathcal{P}_3$}
\label{app:optim}
The stair climbing and the standing up scenarios were validated with the trajectory optimization scheme provided in \citeauthor{Carpentier2016}. 
\deladp{The optimization is formulated as an optimal control problem that aims at generating a valid trajectory for the center of mass.}
To address the other scenarios, we propose a new implementation, entirely open source (\url{https://github.com/stonneau/python_sandbox}), which can be integrated directly in our motion planner software. This formulation uses the center of mass acceleration and angular momentum as input variables, while previously the contact forces were used.
The center-of-mass trajectory resulting from the optimization is then turned into a collision-free whole-body trajectory.

@@

\begin{align} \label{eq:new_eul_acc}
\mathbf{G} \bm{\beta} = 
\underbrace{\mat{m (\mathbf{\ddot{c}} - \mathbf{g}) \\m \mathbf{c} \times (\mathbf{\ddot{c}} - \mathbf{g}) + \dot{L}}}_{\mathbf{w}}
%~ \underbrace{m (\mathbf{\ddot{c}} - \mathbf{g})  \\ m (\mathbf{\ddot{dc}} - \mathbf{g}) + \mathbf{\dot{L}} }_{\mathbf{w}}
\end{align}

where $\dot{L}$ is the angular momentum expressed at the com.
Eq.~\ref{eq:new_eul_acc} defines a 6-dimensional cone $\mathcal{K}$ ~\citep{qiu:dhm:2011,Caron2015}. For a given set of contacts,
this cone determines all the admissible wrenches $\mathbf{w}$ that can be generated by contact forces inside their friction cones.
The face form of $\mathcal{K}$ can be computed using the double description method~\citep{Fukuda1996}, resulting in the following linear inequalities:

@@

The objective 
is then to plan a trajectory for the center of mass such that the generated $\mathbf{w}$ always verifies Eq.~\ref{eq:cone_k}. 
We now consider two contact configurations  $\mathbf{q}_0$ and $\mathbf{q}_1$ computed by our planner: in the general case one contact is broken and one created to get from
$\mathbf{q}_0$ to $\mathbf{q}_1$. We manually define the duration of each of the three contact phases.
In each phase $s$ the centroidal wrench $\mathbf{w}$ is constrained to lie inside a cone $\mathcal{K}_s, s =0 \dots 2$.
We call the total duration of the motion $T$, and formulate the following optimization problem:

\begin{equation} \label{eq:lin_prog_rob} \begin{aligned}
%~ \find \quad & \mx{R} \in \Rv{3\times T}, \mxdd{C} \in  \Rv{6\times T} \\
\minimize_{\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{\ddot{c}}(t), \mathbf{w}(t)}  \quad & \sum_{s=0}^{2} \int_{t_s}^{t_s + \Delta t_s}  h(\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{\ddot{c}}(t), \mathbf{w}(t)) \text{dt}  \\
\st \quad & \mathbf{\ddot{c}}(t) = f(\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{w}(t)) \\
	\quad & \mathbf{A}_s \mathbf{w}(t)  \leq \mathbf{b}_s , \forall t \in  \left[t_s, t_s + \Delta t_s \right[, \forall s \\
	\quad & \mathbf{D}_s \mathbf{c}(t)  \leq \mathbf{d}_s , \forall t \in  \left[t_s, t_s + \Delta t_s \right[, \forall s \\
	\quad & \mathbf{c}(0)  = \mathbf{c}_{\mathbf{q}_0} \\
	\quad & \mathbf{c}(T)  = \mathbf{c}_{\mathbf{q}_1} \\
	\quad & \mathbf{c}(0)  = \mathbf{\dot{c}}(0)  = \mathbf{\ddot{c}}(0) = \mathbf{0}\\
	\quad & \mathbf{c}(T)  = \mathbf{\dot{c}}(T)  = \mathbf{\ddot{c}}(T) = \mathbf{0}\\
    %~ \quad & ({\vct{r}}\tra\vct{r}=1,  \forall t  ?) \\
    %~ \quad & \delta t \sum\limits_1^{T-1}{\vc{r}^1}\tra {\vc{r}^1} \ge t_{transit} \\
    %~ \quad & \mx{A}_{\int} \mxdd{C} \leq \vc{a}_{\int} \\
    %~ \quad & \mx{V}_{\int} \mxd{C} \leq \vc{v}_{\int} \\
    %~ \quad & \mx{B}_{\int}^{1} \mx{C} \leq \vc{b}_{\int} \\
    %~ \quad & [\vc{c}_0\tra, \vcd{c}_0\tra]\tra = \vc{x}_0 \\
    %~ \quad & \vc{c}_T = \vc{c}_T^{input} \\
\end{aligned} \end{equation}

The cost function $h$ is a weighted sum of the angular momentum and center-of-mass acceleration variation over the whole trajectory. 
The system dynamics $f$ ties together the position, velocity and acceleration of the center of mass. 
$\mathbf{D}_s$ and $\mathbf{d}_s$ denote stacked kinematic constraints on the center of mass position, determined by the active contact locations.
The inequalities for each contact are determined in the same way that the reachable workspace is computed in Appendix~\ref{app:rom}, with
the effector serving as root. % (a triangulated object is a polytope, and as such can be described as a set of linear inequalities).

@@

The output of this optimization problem is an admissible center-of-mass trajectory.
To compute the whole body motion, we use a two-step approach.

First, we plan a kinematic motion for the robot, subject to the contact constraints. We also constrain the center of mass to follow
the computed trajectory. This is achieved using a constraint-based RRT planner~\citep{7759083}. As a result we obtain a collision-free whole-body motion.

@@

This motion is then validated in our dynamic simulator~\citep{DelPrete2015b}, using the trajectory as a reference postural task for an inverse-dynamics
controller that accounts for all the robot constraints. This final result is shown in the companion video.

### 2017-01-20 body/contact.tex
Review appendix C

\subsection{Definition of a contact sequence}
In previous contributions\adnote{Add reference here}, a contact plan is defined as a sequence of \deladp{configurations that describe a} quasi-static equilibrium configurations
for each \repadp{contact}{existing} phase \deladp{of the motion}. For instance, a walk cycle would be described by three key configurations: a double-support configuration, a single-support configuration (a contact is broken), and another double-support configuration (a contact is created). 
Our definition of contact plan differs: between two consecutive configurations we allow both a contact break and a contact creation---if they are on the same effector. 
In the previous example, our contact plan would simply consist of the two double-support configurations. 
%However, if the contact broken and the contact created sequentially are not on the same effector, our planner outputs two distinct states.
This representation is sufficient to describe all the contact constraints of the problem. Furthermore it removes the need to have single-support quasi-static configurations as in the example. 
As shown in the companion video, this allows our framework to produce dynamic motions.

>>>>

\subsection{Definition of a contact sequence}
In previous contributions\adnote{Add reference here}, a contact plan is defined as a sequence of quasi-static equilibrium configurations
for each contact phase. For instance, a walk cycle would be described by three key configurations: a double-support configuration, a single-support configuration (a contact is broken), and another double-support configuration (a contact is created). 
Our definition of contact plan differs: between two consecutive configurations we allow both a contact break and a contact creation---if they are on the same effector. 
In the previous example, our contact plan would simply consist of the two double-support configurations. 
%However, if the contact broken and the contact created sequentially are not on the same effector, our planner outputs two distinct states.
This representation is sufficient to describe all the contact phases. Furthermore it removes the need to have single-support quasi-static configurations as in the example. 
As shown in the companion video, this allows our framework to produce dynamic motions.
